---
title: "1st Peer Assesment"
author: "Carlos"
date: "Saturday, March 14, 2015"
output: html_document
---

###Loading Data
This loads the data
```{r}
AMD <- read.csv('repdata-data-activity\\activity.csv')
```

###Mean of Steps per day

```{r}
NumSteps <- tapply(AMD$steps,AMD$date,sum,na.rm=TRUE)
hist(NumSteps)
```

```{r mean}
meanS <-mean(NumSteps)
```

```{r median}
medianS <- median(NumSteps)
```

The mean is :`r meanS`, and the median is : `r medianS`

##Average daily activity pattern

This constructs a time series of the data
```{r}
DailyPattern<- tapply(AMD$steps,AMD$interval,mean,na.rm = TRUE)
plot(DailyPattern,type = 'l' ,col='red',xlab="Intervals",ylab="Average number of steps")
```

It is seen that the maximum average is reach around the 100 interval , we will calculate the exact number with the following code:
```{r}
Index <- which.max(DailyPattern)
Tinterval <- AMD$interval[Index]
```
From which the maximum average of number of steps is reach at the `r Index` interval which comprises the  `r Tinterval`- `r Tinterval - 5` minutes from the beginning of the measurement.


## Imputting missing values

```{r}
Nas <- sum(is.na(AMD$steps))
```
The number of missing values is `r Nas`.  
We will now replace each NAs with the *mean* number of the respective interval in which it occurs.

```{r}
fillin <- function(C,i){
  return(C[[as.character(i)]])
}

newSteps <- AMD$steps
for(i in (1:length(newSteps))){
  t = newSteps[i]
  if(is.na(t)){
    interval = AMD$interval[i]
    newSteps[i]= fillin(DailyPattern,interval)
  }
}
```

Now we create a new data set, updating the *steps* column

```{r}
newData<- AMD
newData$steps <- newSteps
```

We will now recompute the distribution of steps and the mean and median
```{r}
newNumSteps <- tapply(newData$steps,newData$date,sum)
hist(newNumSteps)

newMean <- mean(newNumSteps)
newMedian<- median(newNumSteps)
```

The difference in the mean and median between both data sets is presented in the following table
```{r}
differences<-c(newMean-meanS,newMedian-medianS)
Statistic<-c("mean","median")
df <- data.frame(Statistic,differences)
df
```
A positive change is observed in both cases.

## Differences in activity patterns between weekdays and weekends


This creates a new factor variable with levels *weekday* and *weekend* and add it to the new data set in the variable *day*

```{r}
Days<-weekdays(as.Date(AMD$date))
weekDays<-c("Monday","Tuesday","Wednesday","Thrusday","Friday")

F<-as.factor(!(Days %in% weekDays))

levels(F) <- c("weekday","weekend")

newData$day <- F

WeekdaysData <- newData[newData$day == "weekday",]
WeekendData <- newData[newData$day != "weekday",]
```

We will show to ways of creating the plots, one using base graphics and the other using ggplot2  

###Base graphics
```{r}
par(mfrow = c(2,1) )

AverageStepsWeekdays <- tapply(WeekdaysData$steps,WeekdaysData$interval,mean)
AverageStepsWeekend <- tapply(WeekendData$steps,WeekendData$interval,mean)

plot(AverageStepsWeekdays,type='l',ylab='Average steps',xlab='Intervals')
plot(AverageStepsWeekend,type= 'l',ylab='Average steps',xlab='Intervals')
```

###ggplot2
```{r}
library(ggplot2)

stat_sum_single <- function(fun, geom="point",colour="blue", ...) {
  stat_summary(fun.y=fun, colour=colour, geom=geom, ...)
}


p <- ggplot(newData,aes(interval,steps)) + facet_grid(day~.) + geom_point(alpha = 0.00)
p+ stat_sum_single(mean,geom="line",size=0.5) + coord_cartesian(ylim = c(-10, 300)) + labs(x="Intervals" , y = "Average steps")
```


